Case Study: Gene Set Enrichment Analysis (Mouse Data Set)

### 2018-08-28 datasets/GSEA/fgsea-human.Rmd
Add GSEA analysis; remove benchmarking of goSeq analysis
Adding GSEA analysis (by fgsea package) which implements a FCS
(Functional Class Scoring) method. Unlike goSeq, it does not
require an arbitrary threshold to define DE genes. Instead it
uses the rankings from a DE analysis, and checks for enrichment
of these rankings for all genes. Here the gene set size appears
to be both informative and independent under the null, since
we no longer have to perform selection/filtering of the p-values.

Also removing the benchmarking from the goSeq analyses. This is
because we see that (and logically in

---
title: "Case Study: Gene Set Enrichment Analysis (Human Data Set)"
author: "Alejandro Reyes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

@@

@@

```{r}
library(dplyr)
library(ggplot2)
library(scales)
library(DESeq2)
library(EnsDb.Hsapiens.v86)
library(biomaRt)
library(goseq)

@@

Finally, we’ll perform a gene set enrichment analysis with goseq. goseq performs GSEA and adjusts for gene length biases. The bias is represented in the plot below, with gene length bin on the x-axis and proportion of genes detected to be differentially expressed on the y-axis.

```{r}
if (!file.exists(result_file)) {
    ## get median transcript length for bias adjustment
    txByGene <- transcriptsBy(EnsDb.Hsapiens.v86, "gene")
    geneLength <- sapply(width(txByGene), median)
    geneLength <- geneLength[names(genes)]
    genes[is.na(genes)] <- 0

    ## perform gsea
    pwf <- nullp(genes, bias.data = geneLength)
    goRes <- goseq(pwf, gene2cat = goSets)

    saveRDS(goRes, file = result_file)
} else {
    goRes <- readRDS(result_file)
}

## Add random (uninformative) covariate
set.seed(66778)
goRes$rand_covar <- rnorm(nrow(goRes))

@@

We filter out really small gene sets and those gene sets which have no DE genes
in them (many of these are set to 1).

```{r}
goRes <- goRes %>%
  dplyr::filter(numDEInCat > 0, numInCat > 10 )
```

@@

Here, we can see that the covariant variable is actually informative.

@@

```{r informativeCovariate, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
rank_scatter(dat = goRes, pval = "over_represented_pvalue", 
              covariate = "numInCat", bins = 50, funx = log2, 
              funfill = log10_trans()) +
    ylim(0, 12) +
    ggtitle("Over-represented gene sets") +
    xlab(expression(log[10]~"(# of genes)")) +
    ylab(expression(-log[10]~"(p-value)") ) 
```

@@

```{r assumptions, message=FALSE, fig.width=10, fig.height=3.2}
strat_hist(goRes, pval="over_represented_pvalue", covariate="numInCat", maxy=7)
```

@@

```{r informativeCovariate-rand, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
rank_scatter(dat = goRes, pval = "over_represented_pvalue", 
              covariate = "rand_covar", bins = 50, 
              funfill = log10_trans()) +
    ylim(0, 12) +
    ggtitle("Over-represented gene sets") +
    ylab(expression(-log[10]~"(p-value)") ) 
```

@@

```{r assumptions-rand, message=FALSE, fig.width=10, fig.height=3.2}
strat_hist(goRes, pval="over_represented_pvalue", covariate="rand_covar", maxy=7)
```

@@

```{r setup-benckmark, message=FALSE}
## rename columns and prepare for benchmarking
res <- dplyr:::select(goRes, c("over_represented_pvalue", "numInCat", "rand_covar")) %>%
    dplyr:::rename(pval = over_represented_pvalue, ind_covariate = numInCat)
res$pval <- pmin(res$pval, 1)

>>>>

---
title: "Case Study: Gene Set Enrichment Analysis (Human Data Set)"
author: "Alejandro Reyes and Keegan Korthauer"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

@@

Here we use the `fgsea` Bioconductor package to implement the GSEA method. This is 
a Functional Class Scoring approach, which does not require setting an arbitrary threshold
for Differential Expression, but instead relies on the gene's rank (here we rank by
DESeq2 test statistic).

@@

```{r}
library(dplyr)
library(ggplot2)
library(scales)
library(DESeq2)
library(EnsDb.Hsapiens.v86)
library(biomaRt)
library(fgsea)

@@

Now we use the `fgsea` package to perform the gene set enrichment analysis and 
obtain a enrichment p-value for each pathway.

```{r}
# invert the list so each item is a pathway instead of a gene
goSets <- split(rep(names(goSets), lengths(goSets)), unlist(goSets))
stats <- res$stat
names(stats) <- rownames(res)
stats <- stats[!is.na(stats)]

if (!file.exists(result_file)) {
  goRes <- fgsea(goSets, 
                  stats, 
                  nperm=10000, 
                  maxSize=500,
                  minSize=5)  
  saveRDS(goRes, file = result_file)
} else {
  goRes <- readRDS(result_file)
}
```

Add a random (uninformative covariate) to the dataset.

```{r}
## Add random (uninformative) covariate
set.seed(66778)
goRes$rand_covar <- rnorm(nrow(goRes))

@@

@@

Here, we want to check whether the size of the gene set is actually informative and 
independent under the null.

@@

```{r informativeCovariate, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
rank_scatter(dat = goRes, pval = "pval", 
              covariate = "size", bins = 50, funx = log2, 
              funfill = log10_trans()) +
    ggtitle("Enriched gene sets") +
    xlab(expression(log[10]~"(# of genes)")) +
    ylab(expression(-log[10]~"(p-value)") ) 
```

@@

```{r assumptions, message=FALSE, fig.width=10, fig.height=3.2}
strat_hist(goRes, pval="pval", covariate="size", maxy=11)
```

@@

```{r informativeCovariate-rand, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
rank_scatter(dat = goRes, pval = "pval", 
              covariate = "rand_covar", bins = 50, 
              funfill = log10_trans()) +
    ggtitle("Enriched gene sets") +
    ylab(expression(-log[10]~"(p-value)") ) 
```

@@

```{r assumptions-rand, message=FALSE, fig.width=10, fig.height=3.2}
strat_hist(goRes, pval="pval", covariate="rand_covar", maxy=11)
```

@@

```{r setup-benckmark, message=FALSE}
## rename columns and prepare for benchmarking
res <- dplyr:::select(goRes, c("pval", "size", "rand_covar")) %>%
    dplyr:::rename(ind_covariate = size)

### 2018-08-28 datasets/GSEA/fgsea-mouse.Rmd
Add GSEA analysis; remove benchmarking of goSeq analysis
Adding GSEA analysis (by fgsea package) which implements a FCS
(Functional Class Scoring) method. Unlike goSeq, it does not
require an arbitrary threshold to define DE genes. Instead it
uses the rankings from a DE analysis, and checks for enrichment
of these rankings for all genes. Here the gene set size appears
to be both informative and independent under the null, since
we no longer have to perform selection/filtering of the p-values.

Also removing the benchmarking from the goSeq analyses. This is
because we see that (and logically in

---
title: "Case Study: Gene Set Enrichment Analysis (Mouse Data Set)"
author: "Alejandro Reyes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

@@

```{r}
library(dplyr)
library(ggplot2)
library(scales)
library(DESeq2)
library(EnsDb.Mmusculus.v75)
library(biomaRt)
library(goseq)

@@

**Why no filtering on gene_biotype == "protein_coding" for this data? AR: I think I did this in the other vignette to keep running times short. This shouldn't change the result since anyway protein coding genes are those with GO categories and genes without GO assignments are ignored.**

@@

In order to get a list of genes to test for GSEA, we will test each gene for differential gene expression between hematopoietic stem cells and multipotent progenitors (fraction 1). To do this, we will run DESeq2 to retrieve genes differentially expressed at a FDR of 10%.

@@

```{r goseq, message=FALSE}
if (!file.exists(goset_file)) {
    mart <- useMart("ensembl", "mmusculus_gene_ensembl")
    goSets <- getBM(c("ensembl_gene_id", "go_id"), mart = mart,
                    filters = "ensembl_gene_id", values = rownames(res))
    goSets <- goSets[!nchar(goSets$go_id) == 0, ]
    goSets <- with(goSets, split(go_id, ensembl_gene_id))
    saveRDS(goSets, file = goset_file)
} else {
   goSets <- readRDS(goset_file)
}
```

Now we are ready to perform gene set enrichment analysis using `goseq`.

```{r}
if (!file.exists(result_file)) {
    ## getting median transcript length
    txByGene <- transcriptsBy(EnsDb.Mmusculus.v75, "gene")
    geneLength <- sapply(width(txByGene), median)
    geneLength <- geneLength[names(genes)]
    genes[is.na(genes)] <- 0

    ## perform gsea
    pwf <- nullp(genes, bias.data = geneLength)
    goRes <- goseq(pwf, gene2cat = goSets)

    saveRDS(goRes, file = result_file)
} else {
    goRes <- readRDS(result_file)
}

## Add random (uninformative) covariate
set.seed(7476)
goRes$rand_covar <- rnorm(nrow(goRes))

@@

Here, we can see that the size of the gene set is actually informative.

@@

```{r informativeCovariate, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
rank_scatter(dat = goRes, pval = "over_represented_pvalue", 
             covariate = "numInCat", bins = 50, funx = log2, 
             funfill = log10_trans()) +
    ylim(0, 12) +
    ggtitle("Over-represented gene sets") +
    xlab(expression(log[10]~"(# of genes)")) +
    ylab(expression(-log[10]~"(p-value)")) 
```

@@

```{r assumptions, message=FALSE, fig.width=10, fig.height=3.2}
strat_hist(goRes, pval = "over_represented_pvalue", covariate = "numInCat", maxy=7)
```

@@

```{r assumptions-rand, message=FALSE, fig.width=10, fig.height=3.2}
strat_hist(goRes, pval="over_represented_pvalue", covariate="rand_covar", maxy=10)
```

@@

```{r setup-benckmark, message=FALSE}
## rename columns and prepare for benchmarking
res <- dplyr::select(goRes, over_represented_pvalue, numInCat, rand_covar) %>%
    dplyr::rename(pval = over_represented_pvalue,
                  ind_covariate = numInCat) %>%
    dplyr::mutate(pval = pmin(pval, 1))

>>>>

---
title: "Case Study: Gene Set Enrichment Analysis (Mouse Data Set)"
author: "Alejandro Reyes and Keegan Korthauer"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

@@

```{r}
library(dplyr)
library(ggplot2)
library(scales)
library(DESeq2)
library(EnsDb.Mmusculus.v75)
library(fgsea)

@@

@@

In order to rank the list of genes for GSEA, we will test each gene for 
differential gene expression between hematopoietic stem cells and multipotent 
progenitors (fraction 1). To do this, we will run DESeq2 to obtain a statistic 
for differential expression.

@@

```{r gocat, message=FALSE}
if (!file.exists(goset_file)) {
    mart <- useMart("ensembl", "mmusculus_gene_ensembl")
    goSets <- getBM(c("ensembl_gene_id", "go_id"), mart = mart,
                    filters = "ensembl_gene_id", values = rownames(res))
    goSets <- goSets[!nchar(goSets$go_id) == 0, ]
    goSets <- with(goSets, split(go_id, ensembl_gene_id))
    saveRDS(goSets, file = goset_file)
} else {
   goSets <- readRDS(goset_file)
}
```

Now we use the `fgsea` package to perform the gene set enrichment analysis and 
obtain a enrichment p-value for each pathway.

```{r}
# invert the list so each item is a pathway instead of a gene
goSets <- split(rep(names(goSets), lengths(goSets)), unlist(goSets))
stats <- res$stat
names(stats) <- rownames(res)
stats <- stats[!is.na(stats)]

if (!file.exists(result_file)) {
  goRes <- fgsea(goSets, 
                  stats, 
                  nperm=10000, 
                  maxSize=500,
                  minSize=5)  
  saveRDS(goRes, file = result_file)
} else {
  goRes <- readRDS(result_file)
}
```

Add a random (uninformative covariate) to the dataset.

```{r}
## Add random (uninformative) covariate
set.seed(7476)
goRes$rand_covar <- rnorm(nrow(goRes))

@@

Here, we want to check whether the size of the gene set is actually informative and 
independent under the null.

@@

```{r informativeCovariate, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
rank_scatter(dat = goRes, pval = "pval", 
             covariate = "size", bins = 50, funx = log2, 
             funfill = log10_trans()) +
    ggtitle("Enriched gene sets") +
    xlab(expression(log[10]~"(# of genes)")) +
    ylab(expression(-log[10]~"(p-value)")) 
```

@@

```{r assumptions, message=FALSE, fig.width=10, fig.height=3.2}
strat_hist(goRes, pval = "pval", covariate = "size", maxy=12)
```

@@

```{r assumptions-rand, message=FALSE, fig.width=10, fig.height=3.2}
strat_hist(goRes, pval="pval", covariate="rand_covar", maxy=10)
```

@@

```{r setup-benckmark, message=FALSE}
## rename columns and prepare for benchmarking
res <- dplyr::select(goRes, pval, size, rand_covar) %>%
    dplyr::rename(pval = pval,
                  ind_covariate = size)

### 2018-08-28 datasets/GSEA/goseq-human.Rmd
Add GSEA analysis; remove benchmarking of goSeq analysis
Adding GSEA analysis (by fgsea package) which implements a FCS
(Functional Class Scoring) method. Unlike goSeq, it does not
require an arbitrary threshold to define DE genes. Instead it
uses the rankings from a DE analysis, and checks for enrichment
of these rankings for all genes. Here the gene set size appears
to be both informative and independent under the null, since
we no longer have to perform selection/filtering of the p-values.

Also removing the benchmarking from the goSeq analyses. This is
because we see that (and logically in

@@

## intermediary files we create below
count_file <- file.path(datdir, "human-counts.rds")
deseq_file <- file.path(datdir, "human-deseq.rds")
goset_file <- file.path(datdir, "human-gosets.rds")
result_file <- file.path(resdir, "human-results.rds")
bench_file <- file.path(sbdir, "human-benchmark.rds")
bench_file_uninf <- file.path(sbdir, "human-uninf-benchmark.rds")
```

@@

```{r}
goRes <- goRes %>%
  dplyr::filter(numDEInCat > 0, numInCat > 10 )
```

@@

We can also explore if the covariate seems to be independent under the null.

@@

### Random

We will explore whether the random covariate can be used as a covariate for modern multiple-testing correction methods in the context of GSEA. In the plot below, the log10 of the p-values is plotted as a function of the 
random covariate. This covariate looks independent of the p-values.

```{r informativeCovariate-rand, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
rank_scatter(dat = goRes, pval = "over_represented_pvalue", 
              covariate = "rand_covar", bins = 50, 
              funfill = log10_trans()) +
    ylim(0, 12) +
    ggtitle("Over-represented gene sets") +
    ylab(expression(-log[10]~"(p-value)") ) 
```

We can also explore if the covariate seems to be independent under the null.

```{r assumptions-rand, message=FALSE, fig.width=10, fig.height=3.2}
strat_hist(goRes, pval="over_represented_pvalue", covariate="rand_covar", maxy=7)
```

## Multiple-Testing Correction

We then execute the benchDesign and generate a *SummarizedBenchmark* object containing multiple-testing corrections using several methods.

```{r setup-benckmark, message=FALSE}
## rename columns and prepare for benchmarking
res <- dplyr:::select(goRes, c("over_represented_pvalue", "numInCat", "rand_covar")) %>%
    dplyr:::rename(pval = over_represented_pvalue, ind_covariate = numInCat)
res$pval <- pmin(res$pval, 1)

## generate default BenchDesign
bd <- initializeBenchDesign()
```

We don't include `ashq`, `fdrreg-e` and `fdrreg-t` from the analysis because
the necessary assumptions are not met in the current case study. Namely, effect sizes
and standard errors are not available for ASH, and the test statistics are not normally
distributed under the null and alternative, as required by Scott's FDR regression methods.

```{r run-benchmark}
if (!file.exists(bench_file)) {
  sGSEA <- buildBench(bd, res, ftCols = "ind_covariate")
  saveRDS(sGSEA, file = bench_file)
} else {
  sGSEA <- readRDS(bench_file)
}
```

We'll also compare the results to an uninformative (random) covariate.

```{r run-benchmark-rand}
if (!file.exists(bench_file_uninf)) {
  res$ind_covariate <- res$rand_covar
  sGSEA_rand <- buildBench(bd, res, ftCols = "ind_covariate")
  saveRDS(sGSEA_rand, file = bench_file_uninf)
} else {
  sGSEA_rand <- readRDS(bench_file_uninf)
}
```

## Benchmark Metrics

```{r}
assayNames(sGSEA) <- "qvalue"
sGSEA <- addDefaultMetrics(sGSEA)
sGSEA <- estimatePerformanceMetrics(sGSEA, addColData=TRUE)
```

```{r plottingResults}
rejections_scatter(sGSEA, as_fraction=FALSE, supplementary=FALSE)
```

```{r}
plotFDRMethodsOverlap(sGSEA, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100)
```

```{r}
covariateLinePlot(sGSEA, alpha = 0.1, covname = "ind_covariate", trans = "log1p")
```

### Random

```{r}
assayNames(sGSEA_rand) <- "qvalue"
sGSEA_rand <- addDefaultMetrics(sGSEA_rand)
sGSEA_rand <- estimatePerformanceMetrics(sGSEA_rand, addColData=TRUE)
```

```{r plottingResults-rand}
rejections_scatter(sGSEA_rand, as_fraction=FALSE, supplementary=FALSE)
```

```{r}
plotFDRMethodsOverlap(sGSEA_rand, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100)
```

```{r}
covariateLinePlot(sGSEA_rand, alpha = 0.1, covname = "ind_covariate", trans = "log1p")
```

## Covariate comparison

Here we compare the method ranks for the two covariates at alpha = 0.10.

```{r}
plotMethodRanks(c(bench_file, bench_file_uninf), 
                colLabels = c("Set Size", "Random"), 
                alpha = 0.10, xlab = "Covariate", 
                excludeMethods = NULL)
```

>>>>

Here we use the `goseq` Bioconductor package to implement the gene set analysis. This is 
an Over-Representation Analysis which does requires setting an arbitrary threshold
for Differential Expression. The test concerns whether any of the GO sets are enriched
for DE genes.

@@

## intermediary files we create below
count_file <- file.path(datdir, "human-counts.rds")
deseq_file <- file.path(datdir, "human-deseq.rds")
goset_file <- file.path(datdir, "human-gosets.rds")
result_file <- file.path(resdir, "human-results-goseq.rds")
```

@@

```{r}
goRes <- goRes %>%
  dplyr::filter(numDEInCat > 0, numInCat > 5 )
```

@@

But unfortunately the distribution 
p-values is quite different for the different covariate strata. This is not surprising, as
we expect that smaller gene sets that have at least one DE gene in them will be skewed 
toward smaller p-values (as compared to larger gene sets that have at least one DE gene), 
simply because a single DE gene represents a larger proportion in the smaller sets.
If we don't condition on sets that have at least one DE gene, however, the distribution 
of p-values is spiked at 1 since goseq assigns all such sets to a p-value of 1.

@@

This suggests that
the covariate is *not* independent under the null hypothesis, so the assumptions
of many of the methods (which use an independent covariate) are not satisfied. 
Therefore we will not proceed with the benchmarking using these results.

### 2018-08-28 datasets/GSEA/goseq-mouse.Rmd
Add GSEA analysis; remove benchmarking of goSeq analysis
Adding GSEA analysis (by fgsea package) which implements a FCS
(Functional Class Scoring) method. Unlike goSeq, it does not
require an arbitrary threshold to define DE genes. Instead it
uses the rankings from a DE analysis, and checks for enrichment
of these rankings for all genes. Here the gene set size appears
to be both informative and independent under the null, since
we no longer have to perform selection/filtering of the p-values.

Also removing the benchmarking from the goSeq analyses. This is
because we see that (and logically in

## intermediary files we create below
count_file <- file.path(datdir, "mouse-counts.rds")
deseq_file <- file.path(datdir, "mouse-deseq.rds")
goset_file <- file.path(datdir, "mouse-gosets.rds")
result_file <- file.path(resdir, "mouse-results.rds")
bench_file <- file.path(sbdir, "mouse-benchmark.rds")
bench_file_uninf <- file.path(sbdir, "mouse-uninf-benchmark.rds")
```

@@

And at the same time, it seems to be fair to assume that it is independent under the null hypothesis.

@@

### Random

We will explore whether the random covariate can be used as a covariate for modern multiple-testing correction methods in the context of GSEA. In the plot below, the log10 of the p-values is plotted as a function of the 
random covariate. This covariate looks independent of the p-values.

```{r informativeCovariate-rand, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
rank_scatter(dat = goRes, pval = "over_represented_pvalue", 
              covariate = "rand_covar", bins = 50, 
              funfill = log10_trans()) +
    ylim(0, 12) +
    ggtitle("Over-represented gene sets") +
    ylab(expression(-log[10]~"(p-value)") ) 
```

We can also explore if the covariate seems to be independent under the null.

```{r assumptions-rand, message=FALSE, fig.width=10, fig.height=3.2}
strat_hist(goRes, pval="over_represented_pvalue", covariate="rand_covar", maxy=10)
```

## Multiple-Testing Correction

Generating the *SummarizedBenchmark* object:

```{r setup-benckmark, message=FALSE}
## rename columns and prepare for benchmarking
res <- dplyr::select(goRes, over_represented_pvalue, numInCat, rand_covar) %>%
    dplyr::rename(pval = over_represented_pvalue,
                  ind_covariate = numInCat) %>%
    dplyr::mutate(pval = pmin(pval, 1))

## generate default BenchDesign
bd <- initializeBenchDesign()
```

We don't include `ashq`, `fdrreg-e` and `fdrreg-t` from the analysis because
the necessary assumptions are not met in the current case study. Namely, effect sizes
and standard errors are not available for ASH, and the test statistics are not normally
distributed under the null and alternative, as required by Scott's FDR regression methods.

```{r run-benchmark}
if (!file.exists(bench_file)) {
    sGSEAMouse <- buildBench(bd, res, ftCols = "ind_covariate")
    saveRDS(sGSEAMouse, file = bench_file)
} else {
    sGSEAMouse <- readRDS(bench_file)
}
```

We'll also compare the results to an uninformative (random) covariate.

```{r run-benchmark-rand}
if (!file.exists(bench_file_uninf)) {
  res$ind_covariate <- res$rand_covar
  sGSEAMouse_rand <- buildBench(bd, res, ftCols = "ind_covariate")
  saveRDS(sGSEAMouse_rand, file = bench_file_uninf)
} else {
  sGSEAMouse_rand <- readRDS(bench_file_uninf)
}
```

## Benchmark Metrics

```{r}
assayNames(sGSEAMouse) <- "qvalue"
sGSEAMouse <- addDefaultMetrics(sGSEAMouse)
```

```{r plottingResults}
rejections_scatter(sGSEAMouse, as_fraction = FALSE, supplementary = FALSE)
```

```{r}
plotFDRMethodsOverlap(sGSEAMouse, alpha = 0.1, supplementary = FALSE,
                      order.by = "freq", nsets = 100)
```

```{r}
covariateLinePlot(sGSEAMouse, alpha = 0.1, covname = "ind_covariate", trans="log1p")
```

### Random

```{r}
assayNames(sGSEAMouse_rand) <- "qvalue"
sGSEAMouse_rand <- addDefaultMetrics(sGSEAMouse_rand)
sGSEAMouse_rand <- estimatePerformanceMetrics(sGSEAMouse_rand, addColData=TRUE)
```

```{r plottingResults-rand}
rejections_scatter(sGSEAMouse_rand, as_fraction=FALSE, supplementary=FALSE)
```

```{r}
plotFDRMethodsOverlap(sGSEAMouse_rand, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100)
```

```{r}
covariateLinePlot(sGSEAMouse_rand, alpha = 0.1, covname = "ind_covariate", trans = "log1p")
```

## Covariate comparison

Here we compare the method ranks for the two covariates at alpha = 0.10.

```{r}
plotMethodRanks(c(bench_file, bench_file_uninf), 
                colLabels = c("Set Size", "Random"), 
                alpha = 0.10, xlab = "Covariate", 
                excludeMethods = NULL)
```

>>>>

## intermediary files we create below
count_file <- file.path(datdir, "mouse-counts.rds")
deseq_file <- file.path(datdir, "mouse-deseq.rds")
goset_file <- file.path(datdir, "mouse-gosets.rds")
result_file <- file.path(resdir, "mouse-results-goseq.rds")
```

@@

But unfortunately the distribution 
p-values is quite different for the different covariate strata. This is not surprising, as
we expect that smaller gene sets that have at least one DE gene in them will be skewed 
toward smaller p-values (as compared to larger gene sets that have at least one DE gene), 
simply because a single DE gene represents a larger proportion in the smaller sets.
If we don't condition on sets that have at least one DE gene, however, the distribution 
of p-values is spiked at 1 since goseq assigns all such sets to a p-value of 1.

@@

This suggests that
the covariate is *not* independent under the null hypothesis, so the assumptions
of many of the methods (which use an independent covariate) are not satisfied. 
Therefore we will not proceed with the benchmarking using these results.
